### 2022-03-09 bools/bools.tex
more

In the last decades, automatic theorem provers have been used
successfully as backends to ``hammers'' in proof assistants
\cite{ku-15-holyhammer, pb-12-sh} and to software
verifiers \cite{fp-13-why3}. Most advanced provers, such as CVC4
\cite{cbetal-11-cvc4}, E \cite{scv-19-e23}, and Vampire
\cite{lkav-13-vampire}, are based on first-order logic, whereas
most frontends that use them are based on versions of higher-order logic. Thus,
there is a large gap in expressiveness between front- and backends.
%
\looseness=-1
This gap is bridged using well-known translations from higher-order to
first-order logic \cite{ar-70-hol,
mp-08-trans}. However, as shown in Chapter \ref{ch:ehoh}, translations are usually less efficient than native support. The
distinguishing features of higher-order logic used by proof assistants that the
translation must eliminate include
$\lambda$-binders, function extensionality---the property that functions are
equal if they agree on every argument, described by the axiom $\iforall xy. \allowbreak\, (\iforall z. \, x \, z \ieq y \,z)
\iimplies x \ieq y$, and formulas occurring as arguments of function symbols \cite{mp-08-trans}.

@@

We created two rules that are syntactically similar to \infname{FP} but
are adapted for higher-order logic with one key distinction -- we do not perform any translation:
%
\begin{align*}
  & \namedinference{Cases}{C[f]}{C[\ifalse] \vee f \eq \itrue}
  && \namedsimp{CasesSimp}{C[f]}{C[\ifalse] \vee f \eq \itrue \quad C[\itrue] \vee f \noteq \itrue}
\end{align*}
%
\looseness=-1
Clearly, the prover that uses the rules should not include them both at the same time.
In addition, since
literals $f \eq \ifalse$ are represented as negative equations $f \noteq \itrue$, which cannot be used to paramodulate from,
we change the first requirement on the order to $\ifalse \succ
\itrue$.

@@

\looseness=-1
While experimenting with our implementation we have noticed some proof patterns
that led us to come up with the following modifications. First, it often
suffices to perform \infname{PI} only on initial clauses -- which is why we
allow the rule to be applied only to the clauses created using at most $k$
generating inferences. Second, if the rule was used in the proof, its premise
is usually only used as part of that inference -- which is why we implemented
a version of \infname{PI} that removes the clause after all possible \infname{PI}
inferences have been performed. We observed that the mode \emph{Imit}$_\star$ is useful in practice since often only a single
approximation of a logical symbol is necessary.

@@

Most provers natively support extensionality reasoning: Bhayat et al. \cite{br-20-full-sup-w-combs} modify
first-order unification to return unification constraints consisting of pairs of
terms of functional type, whereas Steen relies on the unification rules of
Leo-III's calculus \cite[Section 4.3.3.]{as-18-phd} to deal with extensionality.
Bentkamp et al \cite{bbtvw-21-sup-lam} altered core generating inference rules of the superposition calculus
to support extensionality. Instead of requiring that terms
involved in the inference are unifiable, it is required that they can be
decomposed into \emph{disagreement pairs} such that at least one of the
disagreement pairs is of functional type. Disagreement pairs of terms $s$ and
$t$ of the same type are defined inductively using function $\textsf{dp}$:
$\textsf{dp}(s,t) = \emptyset$ if $s$ and $t$ are equal; $\textsf{dp}(a \,
\tuplen{s}, b \, \tuple{t}{m}) = \{(a \,
\tuplen{s}, b \, \tuple{t}{m})\}$ if $a$ and $b$ are different heads;
$\textsf{dp}(\lam{x}{s}, \lam{y}{t}) = \{ (\lam{x}{s}, \lam{y}{t}) \}$;
$\textsf{dp}(a \, \tuplen{s}, a \, \tuplen{t}) = \bigcup_{i=1}^{n}
\textsf{dp}(s_i, t_i)$. Then the extensionality rules are stated as follows:
$$
\namedinference{AbsSup}
  {s \eq t \llor C \qquad u[s'] \mathrel{\dot{\eq}} v \llor D}
  {\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor  u[t] \mathrel{\dot{\eq}} v \llor C \llor D}} $$
\pagebreak[2]
\begin{align*}
  &\namedinference{AbsER}{s \noteq s' \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \cdots \llor  C}} \\
  &\namedinference{AbsEF}{s \eq t \llor  s' \eq u \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor t \noteq u \llor  s' \eq u \llor C}}
\end{align*}

@@

\looseness=-1
Rules \infname{Cases} and \infname{CasesSimp} deal with Boolean terms,
but we need to rely on extensionality reasoning to deal with $\lambda$-abstractions
whose body has type $o$. Using the observation that the formula $\iforall \tuplen{x}. \, f$ 
implies that $\lam{\tuplen{x}}{f}$ is extensionally equal to $\lam{\tuplen{x}}{\itrue}$ (and similarly, if
$\iforall \tuplen{x}. \, \inot f$, then $\lam{\tuplen{x}}{f} \eq \lam{\tuplen{x}}{\ifalse})$, we
designed the following rule (where all loosely bound variables of $f$ are $\tuplen{x}$):
%\pagebreak[2]
%
$$ \namedinference{Interpret$\lambda$}{C[\lam{\tuplen{x}}{f}]}
{(\forall {\tuplen{x}}.\, f) \noteq \itrue \llor C[\lam{\tuplen{x}}{\itrue}] \qquad
(\forall {\tuplen{x}}.\, \inot f) \noteq \itrue \llor C[\lam{\tuplen{x}}{\ifalse}]} $$
%

@@

\subsection{Evaluation of the Rules}

\looseness=-1
For this part of the evaluation, we fixed a single well-performing Zipperposition configuration
called \emph{base} (\emph{b}). Since we are testing a single configuration, we
used the CPU time limit of \NumberOK{15} s -- roughly the time a single configuration is
given in a portfolio mode. Configuration \emph{b} uses the pragmatic variant
pv$_{1121}^{2}$ (Sect.~\ref{sec:unif:evaluation}) of the unification algorithm described in Chapter~\ref{ch:unif}. 
It enables \infname{BoolSimp} rule,
\infname{EC} rule, \infname{PI} rule in \emph{Pragmatic} mode with $k=2$,
\infname{ElimLeibniz} and \infname{ElimPredVar} rules, \infname{BoolER} rule, and \infname{BoolEF} rules. 
To evaluate the usefulness of
all rules we described above, we enable, disable or change the
parameters of a single rule, while keeping all other parameters of $b$ intact. 
In figures that contain sufficiently different configurations, cells are of the form $n (m)$ where $n$ is the total
number of proved problems by a particular configuration and $m$ is the number of
unique problems that a given configuration solved, compared to the other
configurations in the same figure. Intersections of rows and columns
denote corresponding combination of parameters.
Result for the base configuration is written in \emph{cursive}; the best result
is written in \textbf{bold}.

@@

\subsection{Comparison with Other Higher-Order Provers}

@@

Leo-III and Satallax are cooperative theorem provers -- they periodically invoke
first-order provers to finish the proof attempt. Leo-III uses CVC4, E and
iProver \cite{kk-08-iprover} as backends, while Satallax uses Ehoh
% \cite{EHOH CHAPTER}
 as backend. Zipperposition can use Ehoh as backend as well. To test how successful each calculus is, we run
the cooperative provers in two versions: \emph{pure}, which disables backends,
and \emph{coop} which uses all supported backends.

@@

\looseness=-1
Established higher-order provers like Leo-III and Satallax perform very
well on TPTP benchmarks; however, recent evaluations (including the one in Sect.~\ref{sec:ehoh:evaluation}) show that on Sledgehammer
problems they are outperformed by translations to first-order logic
\cite{bbtvw-21-sup-lam, %EHOH CHAPTER,
cbetal-11-cvc4}. Those two provers are built from the ground up as
higher-order provers -- treatment of exclusively higher-order issues such as
extensionality or choice is built into them usually using explosive rules. Those
explosive rules might contribute to their suboptimal performance on mostly
first-order Sledgehammer problems.

>>>>

In Chapter~\ref{ch:intro} we motivated our work by the need to bridge the gap
between higher-order frontends and first-order backends. This gap is
traditionally bridged using well-known translations from higher-order to
first-order logic \cite{ar-70-hol, mp-08-trans}. However, as shown in Chapter
\ref{ch:ehoh}, translations are usually less efficient than native support. The
distinguishing features of higher-order logic used by proof assistants that the
translation must eliminate include $\lambda$-binders, function
extensionality---the property that functions are equal if they agree on every
argument, described by the axiom $\iforall xy. \allowbreak\, (\iforall z. \, x
\, z \ieq y \,z) \iimplies x \ieq y$, and formulas occurring as arguments of
function symbols \cite{mp-08-trans}.

@@

We created two rules that are syntactically similar to \infname{FP} but
are adapted for higher-order logic with one key distinction---we do not perform any translation:
%
\begin{align*}
  & \namedinference{Cases}{C[f]}{C[\ifalse] \vee f \eq \itrue}
  && \namedsimp{CasesSimp}{C[f]}{C[\ifalse] \vee f \eq \itrue \quad C[\itrue] \vee f \noteq \itrue}
\end{align*}
%
\looseness=-1
Clearly, the prover that uses the rules should not include them both at the same time.
In addition, since
literals $f \eq \ifalse$ are represented as negative equations $f \noteq \itrue$, which cannot be used to paramodulate from,
we change the first requirement on the order to $\ifalse \succ
\itrue$.

@@

\looseness=-1
While experimenting with our implementation we have noticed some proof patterns
that led us to come up with the following modifications. First, it often
suffices to perform \infname{PI} only on initial clauses---which is why we
allow the rule to be applied only to the clauses created using at most $k$
generating inferences. Second, if the rule was used in the proof, its premise
is usually only used as part of that inference---which is why we implemented
a version of \infname{PI} that removes the clause after all possible \infname{PI}
inferences have been performed. We observed that the mode \emph{Imit}$_\star$ is useful in practice since often only a single
approximation of a logical symbol is necessary.

@@

Most provers natively support extensionality reasoning: Bhayat et al. \cite{br-20-full-sup-w-combs} modify
first-order unification to return unification constraints consisting of pairs of
terms of functional type, whereas Steen relies on the unification rules of
Leo-III's calculus \cite[Section 4.3.3.]{as-18-phd} to deal with extensionality.
Bentkamp et al \cite{bbtvw-21-sup-lam} altered core generating inference rules of the superposition calculus
to support extensionality. Instead of requiring that terms
involved in the inference are unifiable, it is required that they can be
decomposed into \emph{disagreement pairs} such that at least one of the
disagreement pairs is of functional type. Disagreement pairs of terms $s$ and
$t$ of the same type are defined inductively using function $\textsf{dp}$:
$\textsf{dp}(s,t) = \emptyset$ if $s$ and $t$ are equal; $\textsf{dp}(a \,
\tuplen{s}, b \, \tuple{t}{m}) = \{(a \,
\tuplen{s}, b \, \tuple{t}{m})\}$ if $a$ and $b$ are different heads;
$\textsf{dp}(\lam{x}{s}, \lam{y}{t}) = \{ (\lam{x}{s}, \lam{y}{t}) \}$;
$\textsf{dp}(a \, \tuplen{s}, a \, \tuplen{t}) = \bigcup_{i=1}^{n}
\textsf{dp}(s_i, t_i)$. Then the extensionality rules are stated as follows:
\begin{align*}
  & \namedinference{AbsSup}
  {s \eq t \llor C \qquad u[s'] \mathrel{\dot{\eq}} v \llor D}
  {\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor  u[t] \mathrel{\dot{\eq}} v \llor C \llor D}} \\
  &\namedinference{AbsER}{s \noteq s' \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \cdots \llor  C}} \\
  &\namedinference{AbsEF}{s \eq t \llor  s' \eq u \llor C}{\sigmacl{s_1 \noteq s'_1 \llor \cdots \llor s_n \noteq s'_n \llor t \noteq u \llor  s' \eq u \llor C}}
\end{align*}

@@

\looseness=-1
Rules \infname{Cases} and \infname{CasesSimp} deal with Boolean terms,
but we need to rely on extensionality reasoning to deal with $\lambda$-abstractions
whose body has type $o$. Using the observation that the formula $\iforall \tuplen{x}. \, f$ 
implies that $\lam{\tuplen{x}}{f}$ is extensionally equal to $\lam{\tuplen{x}}{\itrue}$ (and similarly, if
$\iforall \tuplen{x}. \, \inot f$, then $\lam{\tuplen{x}}{f} \eq \lam{\tuplen{x}}{\ifalse})$, we
designed the following rule (where all loosely bound variables of $f$ are $\tuplen{x}$):
%\pagebreak[2]
%
$$ \namedinference{Interpret$\lambda$}{C[\lam{\tuplen{x}}{f}]}
{(\iforall {\tuplen{x}}.\, f) \noteq \itrue \llor C[\lam{\tuplen{x}}{\itrue}] \qquad
(\iforall {\tuplen{x}}.\, \inot f) \noteq \itrue \llor C[\lam{\tuplen{x}}{\ifalse}]} $$
%

@@

\ourpara{Evaluation of the Rules}
\looseness=-1
For this part of the evaluation, we fixed a single well-performing Zipperposition configuration
called \emph{base} (\emph{b}). Since we are testing a single configuration, we
used the CPU time limit of \NumberOK{15} s---roughly the time a single configuration is
given in a portfolio mode. Configuration \emph{b} uses the pragmatic variant
pv$_{1121}^{2}$ (Sect.~\ref{sec:unif:evaluation}) of the unification algorithm described in Chapter~\ref{ch:unif}. 
It enables \infname{BoolSimp} rule,
\infname{EC} rule, \infname{PI} rule in \emph{Pragmatic} mode with $k=2$,
\infname{ElimLeibniz} and \infname{ElimPredVar} rules, \infname{BoolER} rule, and \infname{BoolEF} rules. 
To evaluate the usefulness of
all rules we described above, we enable, disable or change the
parameters of a single rule, while keeping all other parameters of $b$ intact. 
In figures that contain sufficiently different configurations, cells are of the form $n (m)$ where $n$ is the total
number of proved problems by a particular configuration and $m$ is the number of
unique problems that a given configuration solved, compared to the other
configurations in the same figure. Intersections of rows and columns
denote corresponding combination of parameters.
Result for the base configuration is written in \emph{cursive}; the best result
is written in \textbf{bold}.

@@

\ourpara{Comparison with Other Higher-Order Provers}

@@

Leo-III and Satallax are cooperative theorem provers---they periodically invoke
first-order provers to finish the proof attempt. Leo-III uses CVC4, E and
iProver \cite{kk-08-iprover} as backends, while Satallax uses Ehoh
% \cite{EHOH CHAPTER}
 as backend. Zipperposition can use Ehoh as backend as well. To test how successful each calculus is, we run
the cooperative provers in two versions: \emph{pure}, which disables backends,
and \emph{coop} which uses all supported backends.

@@

\looseness=-1
Established higher-order provers like Leo-III and Satallax perform very
well on TPTP benchmarks; however, recent evaluations (including the one in Sect.~\ref{sec:ehoh:evaluation}) show that on Sledgehammer
problems they are outperformed by translations to first-order logic
\cite{bbtvw-21-sup-lam, %EHOH CHAPTER,
cbetal-11-cvc4}. Those two provers are built from the ground up as
higher-order provers---treatment of exclusively higher-order issues such as
extensionality or choice is built into them usually using explosive rules. Those
explosive rules might contribute to their suboptimal performance on mostly
first-order Sledgehammer problems.

### 2022-03-09 unif/unif.tex
more

    \begin{description}
      \item[Deletion] A decreases.
      \item[Decomposition] A decreases.
      \item[Failure] Trivial -- represents a terminal node.
      \item[Solution] This transition applies on 
        constraints of the form
        $F \unif t$. This reduces A, since the constraint $F \unif t$ is removed
        and $F$ cannot appear on the right-hand side.
      \item[Imitation] The rule is applicable only on 
        $\lambda \overline{x}.\, F \, \overline{s}_m  \unif \lambda
        \overline{x}.\, \cst{f} \, \overline{t}_n$. The \textsf{Imitation} transition
        replaces the constraint with $\{H_i \,
        \overline{s}_m  \unif \overline{t}_i \mid 1\le i\le n\}$. This reduces A, as $F$ does not appear on any right-hand side.
    \item[Projection] The rule is applicable only on 
        $\lambda \overline{x}.\, F \, \overline{s}_m  \unif \lambda
        \overline{x}.\, a \, \overline{t}_n$. If $a$ is a bound variable,
        and we project $F$ to argument $s_i$ different from $a$, \textsf{Failure} applies and
        PT trivially terminates. If we project $F$ to $a$, we apply \textsf{Decomposition},
        which reduces $A$. If $a$ is a constant, for \textsf{Failure} not to apply, we
        have to project to a base-type ground term $s_i$. This does not increase A,
        since no variables appear on right-hand sides, but removes the variable $F$
        from $E$, reducing B by one. \qedhere
    \end{description}
\end{proof}

@@

%@PETAR: I moved this example up by one paragraph. 
% Unfortunately, I had to inline the unifier \sigma_3 to make this fit.
% Maybe it should also use the same names for substitutions as in the previous para?
 \begin{exa}
   \label{unif:example:solid}
   \looseness=-1
  Let $\{F \, (\cst{f} \, \cst{a}) \unif \cst{g} \, \cst{a} \, (G \, \cst{a})\}$ be
  the unification problem to solve. Projecting~$F$ onto the first argument will
  lead to a nonunifiable problem, so we perform imitation of~$\cst{g}$ building a
  binding $\sigma_1 = \{ F \mapsto \lambda x.\, \cst{g} \, (F_1 \, x) \, (F_2 \, x)
  \}$. After decomposition, this yields the  problem $\{F_1
  \, (\cst{f} \, \cst{a}) \unif \cst{a}, \allowbreak F_2 \, (\cst{f} \, \cst{a}) \unif G \,
  \cst{a}\}$. Again, we can only imitate $\cst{a}$ for $F_1$ -- building a new 
  binding $\sigma_2 = \{F_1 \mapsto \lambda x.\, \cst{a} \}$. Finally, this
  yields the problem $\{ F_2 \, (\cst{f} \, \cst{a}) \unif G \, \cst{a} \}$.
  According to Lemma \ref{unif:lemma:solid-vars-diff-mgu}, we find CSUs for the
  problems $J_1 \, \cst{a} = \cst{f} \, \cst{a}$ and $I_1 \, (\cst{f} \, \cst{a})
  \unif \cst{a}$ using PT. The latter problem has a singleton CSU $\{I_1 \mapsto \lambda x.\,
  \cst{a}\}$, whereas the former has a CSU containing $\{ J_1 \mapsto \lambda x.\,
  \cst{f} \, x \}$ and $\{ J_1 \mapsto \lambda x.\, \cst{f} \, \cst{a} \}$.
  Combining these solutions, we obtain an MGU $\sigma_3 =
  \{ F_2 \mapsto \lambda x.\, H \, x \, x \, \cst{a},\,
  G   \mapsto \lambda x.\, H \, (\cst{f} \, \cst{a}) \, (\cst{f} \, x) \, x \}$ 
  for $F_2 \, (\cst{f}\, \cst{a}) \unif G \, \cst{a}$.
  %\begin{equation*}
  %    F_2 \mapsto \lambda x.\, H \, x \, x \, \cst{a} \\
  %    G   \mapsto \lambda x.\, H \, (\cst{f} \, \cst{a}) \, (\cst{f} \, x) \, x
  %\end{equation*}
  Finally, we get to the MGU of the original problem $\sigma = \sigma_3 \sigma_2 \sigma_1 = \{ F
  \mapsto \lambda x.\, \cst{g} \, \cst{a} \, (H \, x \, x \, \cst{a}),\allowbreak G \mapsto
  \lambda x.\, H \, (\cst{f} \, \cst{a}) \, (\cst{f} \, x) \, x  \}$.
  For brevity, we omitted the intermediate bindings of auxiliary variables
  in $\sigma$.
\end{exa}

@@

The fundamental idea of first-order fingerprint indexing carries over to
higher-order terms -- 
application of a substitution does not change the rigid skeleton of a term.
However, to extend fingerprint indexing to
higher-order terms, we must address the issues of $\alpha\beta\eta$-normalization and figure out how to cope with $\lambda$-abstractions and
bound variables. To that end, we define a function $\hotofo{t}$, defined on
%TODO: @PETAR \eta-long?
% Yes
$\eta$-long $\beta$-reduced terms in De Bruijn \cite{db-75-dbindices} notation:
%TODO: @PETAR: also mention that this is an encoding into FOL before the definition?
% @ALEX: There is a sentence about that later. I do not see how it can easily fit in this description :(
\begin{align*}
& \hotofo{F \, \overline{s}} = F
&& \hotofo{\dbvar{i} \, \overline{s}_n } = \cstdba{i}(\hotofo{s_1}, \ldots, \hotofo{s_n})
&& \hotofo{\cst{f} \, \overline{s}_n } = \cst{f}(\hotofo{s_1}, \ldots, \hotofo{s_n}) &&
\hotofo{\lambda \overline{x}. \, s  } = \hotofo{s}
\end{align*}
\looseness=-1
We let $\dbvar{i}$ be a bound variable of type $\alpha$ with De Bruijn index $i$,
and $\cstdba{i}$ be a fresh constant corresponding to this variable. All constants
$\cstdba{i}$ must be fresh. 
Effectively, $\hotofo{\;}$ transforms an $\eta$-long $\beta$-reduced higher-order term to an untyped first-order term. Let $\nf{t}{\beta\eta}$ be
the $\eta$-long $\beta$-reduced form of $t$; the higher-order generic
fingerprinting function $\text{gfpf}_\textsf{ho}$, which relies on conversion $\nametodb{t}$ from
named to De Bruijn representation, is defined as
\[
  \text{gfpf}_\textsf{ho}(t,p) =
    \text{gfpf}(\hotofo{\nametodb{\nf{t}{\beta\eta}}}, p)
\]

@@

The configuration of our procedure with no oracles outperforms the JP procedure with the first-order oracle. This
suggests that the design of the procedure, in particular lazy normalization and
lazy application of the substitution, already reduces the effects of the JP
procedure's main bottlenecks. The raw evaluation data show that on TPTP benchmarks, complete and pragmatic configurations
differ in the set of problems they solve -- cv solves 19
problems not solved by pv$^{4}_{2222}$, whereas pv$^{4}_{2222}$ solves 34
problems cv does not solve. Similarly, comparing the pragmatic configurations with each other,
pv$^{6}_{3333}$ and pv$^{4}_{2222}$ each solve 13 problems that the other one does not.  The overall higher success rate of pv$^2_{1020}$ compared to
pv$^2_{1222}$ suggests that solving flex-flex pairs by trivial unifiers often
suffices for superposition-based theorem proving.

@@

% TODO: Maybe investigate at some point in future.
% An interesting observation is that including the solid oracle has a different
% effect on TPTP and Sledgehammer problems. However, solid fragment does not
% subsume the first-order fragment -- first-order terms that share variables or
% are not linear are not solid. This can also explain why solid unification helps
% on TPTP problems that usually include hard hand-crafted higher-order problems,
% but hurts on mostly first-order Sledgehammer problems.
\looseness=-1 
A subset of TPTP benchmarks, concerning operations on Church numerals, is designed to test the efficiency of
higher-order unification. Our procedure performs exceptionally well
on these problems -- it solves all of them, usually
faster than other competitive higher-order provers.
  There are 11 benchmarks in the \texttt{NUM} category of TPTP that contain conjectures about
  Church numerals: \texttt{NUM020\^{}1}, \texttt{NUM021\^{}1}, \texttt{NUM415\^{}1},
  \texttt{NUM416\^{}1}, \texttt{NUM417\^{}1}, \texttt{NUM418\^{}1}, \texttt{NUM419\^{}1}, \texttt{NUM798\^{}1},
  \texttt{NUM799\^{}1}, \texttt{NUM800\^{}1}, and \texttt{NUM801\^{}1}.
  We evaluated those problems using the same CPU nodes and the same time limits 
  as above. In addition to Zipperposition, we used
  all higher-order provers that took part in the 2019 edition of CASC \cite{gs-19-casc27} (in the THF category) for this evaluation: CVC4
  1.7 \cite{cbetal-11-cvc4}, Leo-III 1.4 \cite{sb-21-leo3}, Satallax 3.4
  \cite{cb-12-satallax}, Vampire 4.4 THF \cite{lkav-13-vampire}. Figure~\ref{fig:numerals}
  shows the CPU time needed to solve a problem or ``--'' if the prover timed out.
  \pagebreak[2]

>>>>

    \begin{description}
      \item[Deletion] A decreases.
      \item[Decomposition] A decreases.
      \item[Failure] Trivial---represents a terminal node.
      \item[Solution] This transition applies on 
        constraints of the form
        $F \unif t$. This reduces A, since the constraint $F \unif t$ is removed
        and $F$ cannot appear on the right-hand side.
      \item[Imitation] The rule is applicable only on 
        $\lambda \overline{x}.\, F \, \overline{s}_m  \unif \lambda
        \overline{x}.\, \cst{f} \, \overline{t}_n$. The \textsf{Imitation} transition
        replaces the constraint with $\{H_i \,
        \overline{s}_m  \unif \overline{t}_i \mid 1\le i\le n\}$. This reduces A, as $F$ does not appear on any right-hand side.
    \item[Projection] The rule is applicable only on 
        $\lambda \overline{x}.\, F \, \overline{s}_m  \unif \lambda
        \overline{x}.\, a \, \overline{t}_n$. If $a$ is a bound variable,
        and we project $F$ to argument $s_i$ different from $a$, \textsf{Failure} applies and
        PT trivially terminates. If we project $F$ to $a$, we apply \textsf{Decomposition},
        which reduces $A$. If $a$ is a constant, for \textsf{Failure} not to apply, we
        have to project to a base-type ground term $s_i$. This does not increase A,
        since no variables appear on right-hand sides, but removes the variable $F$
        from $E$, reducing B by one. \qedhere
    \end{description}
\end{proof}

@@

%@PETAR: I moved this example up by one paragraph. 
% Unfortunately, I had to inline the unifier \sigma_3 to make this fit.
% Maybe it should also use the same names for substitutions as in the previous para?
 \begin{exa}
   \label{unif:example:solid}
   \looseness=-1
  Let $\{F \, (\cst{f} \, \cst{a}) \unif \cst{g} \, \cst{a} \, (G \, \cst{a})\}$ be
  the unification problem to solve. Projecting~$F$ onto the first argument will
  lead to a nonunifiable problem, so we perform imitation of~$\cst{g}$ building a
  binding $\sigma_1 = \{ F \mapsto \lambda x.\, \cst{g} \, (F_1 \, x) \, (F_2 \, x)
  \}$. After decomposition, this yields the  problem $\{F_1
  \, (\cst{f} \, \cst{a}) \unif \cst{a}, \allowbreak F_2 \, (\cst{f} \, \cst{a}) \unif G \,
  \cst{a}\}$. Again, we can only imitate $\cst{a}$ for $F_1$---building a new 
  binding $\sigma_2 = \{F_1 \mapsto \lambda x.\, \cst{a} \}$. Finally, this
  yields the problem $\{ F_2 \, (\cst{f} \, \cst{a}) \unif G \, \cst{a} \}$.
  According to Lemma \ref{unif:lemma:solid-vars-diff-mgu}, we find CSUs for the
  problems $J_1 \, \cst{a} = \cst{f} \, \cst{a}$ and $I_1 \, (\cst{f} \, \cst{a})
  \unif \cst{a}$ using PT. The latter problem has a singleton CSU $\{I_1 \mapsto \lambda x.\,
  \cst{a}\}$, whereas the former has a CSU containing $\{ J_1 \mapsto \lambda x.\,
  \cst{f} \, x \}$ and $\{ J_1 \mapsto \lambda x.\, \cst{f} \, \cst{a} \}$.
  Combining these solutions, we obtain an MGU $\sigma_3 =
  \{ F_2 \mapsto \lambda x.\, H \, x \, x \, \cst{a},\,
  G   \mapsto \lambda x.\, H \, (\cst{f} \, \cst{a}) \, (\cst{f} \, x) \, x \}$ 
  for $F_2 \, (\cst{f}\, \cst{a}) \unif G \, \cst{a}$.
  %\begin{equation*}
  %    F_2 \mapsto \lambda x.\, H \, x \, x \, \cst{a} \\
  %    G   \mapsto \lambda x.\, H \, (\cst{f} \, \cst{a}) \, (\cst{f} \, x) \, x
  %\end{equation*}
  Finally, we get to the MGU of the original problem $\sigma = \sigma_3 \sigma_2 \sigma_1 = \{ F
  \mapsto \lambda x.\, \cst{g} \, \cst{a} \, (H \, x \, x \, \cst{a}),\allowbreak G \mapsto
  \lambda x.\, H \, (\cst{f} \, \cst{a}) \, (\cst{f} \, x) \, x  \}$.
  For brevity, we omitted the intermediate bindings of auxiliary variables
  in $\sigma$.
\end{exa}

@@

The fundamental idea of first-order fingerprint indexing carries over to
higher-order terms---application of a substitution does not change the rigid skeleton of a term.
However, to extend fingerprint indexing to
higher-order terms, we must address the issues of $\alpha\beta\eta$-normalization and figure out how to cope with $\lambda$-abstractions and
bound variables. To that end, we define a function $\hotofo{t}$, defined on
%TODO: @PETAR \eta-long?
% Yes
$\eta$-long $\beta$-reduced terms in De Bruijn \cite{db-75-dbindices} notation:
%TODO: @PETAR: also mention that this is an encoding into FOL before the definition?
% @ALEX: There is a sentence about that later. I do not see how it can easily fit in this description :(
\begin{align*}
& \hotofo{F \, \overline{s}} = F
&& \hotofo{\dbvar{i} \, \overline{s}_n } = \cstdba{i}(\hotofo{s_1}, \ldots, \hotofo{s_n})
&& \hotofo{\cst{f} \, \overline{s}_n } = \cst{f}(\hotofo{s_1}, \ldots, \hotofo{s_n}) &&
\hotofo{\lambda \overline{x}. \, s  } = \hotofo{s}
\end{align*}
\looseness=-1
We let $\dbvar{i}$ be a bound variable of type $\alpha$ with De Bruijn index $i$,
and $\cstdba{i}$ be a fresh constant corresponding to this variable. All constants
$\cstdba{i}$ must be fresh. 
Effectively, $\hotofo{\;}$ transforms an $\eta$-long $\beta$-reduced higher-order term to an untyped first-order term. Let $\nf{t}{\beta\eta}$ be
the $\eta$-long $\beta$-reduced form of $t$; the higher-order generic
fingerprinting function $\text{gfpf}_\textsf{ho}$, which relies on conversion $\nametodb{t}$ from
named to De Bruijn representation, is defined as
\[
  \text{gfpf}_\textsf{ho}(t,p) =
    \text{gfpf}(\hotofo{\nametodb{\nf{t}{\beta\eta}}}, p)
\]

@@

The configuration of our procedure with no oracles outperforms the JP procedure with the first-order oracle. This
suggests that the design of the procedure, in particular lazy normalization and
lazy application of the substitution, already reduces the effects of the JP
procedure's main bottlenecks. The raw evaluation data show that on TPTP benchmarks, complete and pragmatic configurations
differ in the set of problems they solve---cv solves 19
problems not solved by pv$^{4}_{2222}$, whereas pv$^{4}_{2222}$ solves 34
problems cv does not solve. Similarly, comparing the pragmatic configurations with each other,
pv$^{6}_{3333}$ and pv$^{4}_{2222}$ each solve 13 problems that the other one does not.  The overall higher success rate of pv$^2_{1020}$ compared to
pv$^2_{1222}$ suggests that solving flex-flex pairs by trivial unifiers often
suffices for superposition-based theorem proving.

@@

% TODO: Maybe investigate at some point in future.
% An interesting observation is that including the solid oracle has a different
% effect on TPTP and Sledgehammer problems. However, solid fragment does not
% subsume the first-order fragment -- first-order terms that share variables or
% are not linear are not solid. This can also explain why solid unification helps
% on TPTP problems that usually include hard hand-crafted higher-order problems,
% but hurts on mostly first-order Sledgehammer problems.
\looseness=-1 
A subset of TPTP benchmarks, concerning operations on Church numerals, is designed to test the efficiency of
higher-order unification. Our procedure performs exceptionally well
on these problems---it solves all of them, usually
faster than other competitive higher-order provers.
  There are 11 benchmarks in the \texttt{NUM} category of TPTP that contain conjectures about
  Church numerals: \texttt{NUM020\^{}1}, \texttt{NUM021\^{}1}, \texttt{NUM415\^{}1},
  \texttt{NUM416\^{}1}, \texttt{NUM417\^{}1}, \texttt{NUM418\^{}1}, \texttt{NUM419\^{}1}, \texttt{NUM798\^{}1},
  \texttt{NUM799\^{}1}, \texttt{NUM800\^{}1}, and \texttt{NUM801\^{}1}.
  We evaluated those problems using the same CPU nodes and the same time limits 
  as above. In addition to Zipperposition, we used
  all higher-order provers that took part in the 2019 edition of CASC \cite{gs-19-casc27} (in the THF category) for this evaluation: CVC4
  1.7 \cite{cbetal-11-cvc4}, Leo-III 1.4 \cite{sb-21-leo3}, Satallax 3.4
  \cite{cb-12-satallax}, Vampire 4.4 THF \cite{lkav-13-vampire}. Figure~\ref{fig:numerals}
  shows the CPU time needed to solve a problem or ``--'' if the prover timed out.
  \pagebreak[2]
